Revisiting the Austral Spring Extratropical Southern Hemisphere zonally asymmetric circulation using complex Empirical Orthogonal Functions

### 2022-05-27 analysis/paper/shceof.Rmd
Cambia a british 🧐

Another extensively used approach to characterize the SH tropospheric circulation anomalies, is by computing Empirical Orthogonal Functions (EOF, also known as Principal Component Analysis).
Within the EOF framework, the Southern Annular Mode (SAM) appears as the leading mode of variability of the SH circulation [@fogt2020].
SAM represents a relatively zonally symmetric pattern of alternating low pressures in polar latitude and a ring of high pressures in high latitudes with an embedded wave 3 pattern that is more prominent in the Pacific sector.
The 2nd and 3rd EOFs, usually known as Pacific--South American Patterns (PSA) 1 and PSA2 patterns, respectively, describe meridionally propagating wave trains that originate in the eastern equatorial Pacific and Australian-Indian Ocean sector, and travel towards the South Atlantic following a great-circle arch along the Antarctic coast [@mo2001].
These patterns influence precipitation anomalies in South America [@mo2001].
Although these patterns are usually derived by applying EOF to temporal anomalies, @raphael2003 also applied EOF methods specifically to zonal anomalies.
More recently, @irving2016 proposed a novel methodology for objectively identifying the PSA pattern using Fourier decomposition.

@@

Preliminary analysis showed that the first cEOF was closely related to the zonal wave 1 of Total Ozone Column and the second cEOF was closely related to ENSO.
Therefore, to aid in the interpretation we chose the phase of cEOF1 so that the time series corresponding to the real part has the maximum correlation with the zonal wave 1 of Total Ozone Column between `r knitr::combine_words(LatLabel(o3wave_lats))`.
Similarly, we chose the phase of cEOF2 so that the coefficient of determination between the Oceanic Niño Index [@bamston1997] and the real part is minimized, which also nearly maximizes the correlation with the imaginary part.

@@

Over South America, the Imaginary cEOF2 has positive correlations with precipitation in South Eastern South America (SESA) and central Chile, and negative correlations in eastern Brazil. This correlation field  matches the springtime precipitation signature of ENSO [e.g. @cai2020a] and it is also similar to the precipitation anomalies associated with the A-SAM [@@campitelli2022].

>>>>

Another extensively used approach to characterise the SH tropospheric circulation anomalies, is by computing Empirical Orthogonal Functions (EOF, also known as Principal Component Analysis).
Within the EOF framework, the Southern Annular Mode (SAM) appears as the leading mode of variability of the SH circulation [@fogt2020].
SAM represents a relatively zonally symmetric pattern of alternating low pressures in polar latitude and a ring of high pressures in high latitudes with an embedded wave 3 pattern that is more prominent in the Pacific sector.
The 2nd and 3rd EOFs, usually known as Pacific--South American Patterns (PSA) 1 and PSA2 patterns, respectively, describe meridionally propagating wave trains that originate in the eastern equatorial Pacific and Australian-Indian Ocean sector, and travel towards the South Atlantic following a great-circle arch along the Antarctic coast [@mo2001].
These patterns influence precipitation anomalies in South America [@mo2001].
Although these patterns are usually derived by applying EOF to temporal anomalies, @raphael2003 also applied EOF methods specifically to zonal anomalies.
More recently, @irving2016 proposed a novel methodology for objectively identifying the PSA pattern using Fourier decomposition.

@@

Preliminary analysis showed that the first cEOF was closely related to the zonal wave 1 of Total Ozone Column and the second cEOF was closely related to ENSO.
Therefore, to aid in the interpretation we chose the phase of cEOF1 so that the time series corresponding to the real part has the maximum correlation with the zonal wave 1 of Total Ozone Column between `r knitr::combine_words(LatLabel(o3wave_lats))`.
Similarly, we chose the phase of cEOF2 so that the coefficient of determination between the Oceanic Niño Index [@bamston1997] and the real part is minimised, which also nearly maximises the correlation with the imaginary part.

@@

Over South America, the Imaginary cEOF2 has positive correlations with precipitation in South Eastern South America (SESA) and central Chile, and negative correlations in eastern Brazil. This correlation field  matches the springtime precipitation signature of ENSO [e.g. @cai2020a] and it is also similar to the precipitation anomalies associated with the A-SAM [@campitelli2022].
